---
title: "Reproducible Research project1"
author: "Don Resnik"
date: "May 12, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Loading and preprocessing the data
##What is mean total number of steps taken per day?
```{r}
library(dplyr)
dataValue <- read.csv("activity.csv")
dataValueNoNA <- filter(dataValue,!is.na(dataValue$steps))
#generate histogram of steps per day
stepsPerDay <- aggregate(steps ~ date, dataValueNoNA, sum)
hist(stepsPerDay$steps, main = "Total number of steps per day", xlab = "Steps per day")
# calculate mean and median values
stepsPerDayMean <- mean(stepsPerDay$steps)
stepsPerDayMedian <- median(stepsPerDay$steps)
```

###Steps per day Mean:  `r as.integer(stepsPerDayMean)`
###Steps per day Median:  `r as.integer(stepsPerDayMedian)`

##What is the average daily activity pattern?
```{r}
averageStepsByInterval <- tapply(dataValueNoNA$steps, dataValueNoNA$interval, FUN=mean)
averageStepsByIntervaldf <- data.frame(averageStepsByInterval)
plot(names(averageStepsByInterval), averageStepsByInterval, type = "l", xlab = "5-minute Interval", 
    ylab = "Average number of steps taken", main = "Avg number of steps across 5-minute interval")

# Find the index and the maximum value
maxValueAndInterval <- averageStepsByIntervaldf[which.max(averageStepsByIntervaldf$averageStepsByInterval),]
maxValue <- maxValueAndInterval[[1]]
maxInterval <- names(maxValueAndInterval)
```
###maxInterval `r as.integer(maxInterval)`
###maxValue `r maxValue`

##Imputing missing values
```{r}
# Find NA Values
NAvalues <- filter(dataValue,is.na(dataValue$steps))
numNARows <- nrow(NAvalues)
head(NAvalues)
```

### Number of rows wih NA `r numNARows`

##Fill in all of the missing values in the dataset with the mean for that 5-minute interval.
```{r}
for (i in 1:nrow(NAvalues)) {
    NAvalues[i, 1] <- averageStepsByInterval[[toString(NAvalues[i, 3])]]
}
head(NAvalues)

filledInData <- dataValue
filledInData[is.na(dataValue), 1] <- NAvalues[, 1]
stepsPerDayWithFilledNA <- aggregate(steps ~ date, filledInData, sum)
hist(stepsPerDayWithFilledNA$steps, main = "Total number of steps per day with filled in NA values", xlab = "Steps per day")

# calculate mean and median values
stepsPerDayMeanWithFilledNA <- trunc(mean(stepsPerDayWithFilledNA$steps), digits=0)
stepsPerDayMedianWithFilledNA <- trunc(median(stepsPerDayWithFilledNA$steps), digits=0)
```

###Steps per day with filled in NA Mean:  `r as.integer(stepsPerDayMeanWithFilledNA)`
###Steps per day with filled in NA Median:  `r as.integer(stepsPerDayMedianWithFilledNA)`

### For this case, filling in the NA values did not make much difference in the mean or median number of steps per day.


##Are there differences in activity patterns between weekdays and weekends?
```{r}
filledInData$day <- "weekday"
filledInData$day[weekdays(as.Date(filledInData$date), abb=T) %in% c("Sat","Sun")] <- "weekend"
```


```{r}
library(lattice)
meanStepsPerIntervalWithFilledInData <- aggregate(steps ~ interval + day, data=filledInData, FUN="mean")
xyplot(steps ~ interval | day, data=meanStepsPerIntervalWithFilledInData, type="l", grid=T, layout=c(1,2), ylab="Number of steps", xlab="5-min. intervals", main="Average  5-min. activity for Weekdays and Weekends")
```